SUPPORTING INFORMATION \\ Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-05-12 SI.tex
Andy edits

% Fig. S1
\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=0.9\textwidth, 
        scale=0.5]{SIFig1.pdf}
        \caption{Five-fold cross-validation of the SRV VAMP-2 scores to select the optimal number of SRV coordinates for each sequence. A knee in the VAMP-2 plot was identified at approximately the fifth, fourth, third, and second slow modes for AT-all, GC-end, GC-core, and GC-mix, respectively. An embedding of corresponding dimensionality was then used to cluster frames into discrete states. The absence of any significant separation in the training and testing VAMP-2 scores demonstrates that model is not overfitted.}
        \label{fig:SIFig1}
	\end{center}
\end{figure}

@@

% Fig. S2
\begin{figure}[ht!]
	\centering
    \includegraphics[width=0.7\textwidth, scale=0.5]{SIFig2.pdf}
    \caption{Convergence of the MSM implied time scales $t_i$ as a function of lag time $\tau$. Solid lines indicate maximum likelihood result while dashed lines show the Bayesian ensemble means. The implied time scales for all sequences converge at a lag time of $\tau$ = 1.2 ns (vertical line). The black solid curve marks equality of the implied time scale and lag time and delimits the shaded region wherein the implied time scales are shorter than the lag time and cannot be resolved.}
    \label{fig:SIFig2}
\end{figure}

@@

% Fig. S3
\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=1.1\textwidth]{SIFig3.pdf}
        \caption{Chapman-Kolmogorov (CK) tests comparing the probabilities of remaining within each macrostates for each sequence as a function of lag time predicted by an MSM constructed at the $\tau$ = 1.2 ns lag time (dashed blue line) versus those computed from an MSM constructed at the particular lag time (solid black line). The good agreement between these two results provides numerical validation of the Markovian nature of the $\tau$ = 1.2 ns lag time MSM.}
        \label{fig:SIFig3}
	\end{center}
\end{figure}

@@

% Fig. S4
\begin{figure}[ht!]
	\centering
    \includegraphics[width=.8\textwidth]{SIFig4.pdf}
    \caption{Exponential fits for both slow (dissociation) and fast (fraying) response for all four sequences during ``computational T-jump'' experiments. From 120 independent 1 $\mu$s simulations, we compiled the distribution of times at which terminal A:T and central base pairs first separate to a distance of of 1.3 nm. We extracted our computational estimate of $k_d^\mathrm{fast}$ by fitting a decaying exponential to the fraction of bound A:T termini as a function of time $f_\mathrm{unfrayed}(t) = \exp(-k_d^\mathrm{fast}t)$. Similarly, we extracted our computational estimate of $k_d^\mathrm{slow}$ by fitting a decaying exponential to the fraction of hybridized sequences as a function of time $f_\mathrm{hybridized}(t) = \exp(-k_d^\mathrm{slow}t)$.}
    \label{fig:SIFig4}
\end{figure}

@@

% Fig. S5
\begin{figure}[ht!]
	\centering
    \includegraphics[width=0.8\textwidth]{SIFig5a.pdf}
\end{figure}
\begin{figure}[ht!]
	\centering
    \includegraphics[width=0.8\textwidth]{SIFig5b.pdf}
\end{figure}

@@

The NN model from SantaLucia assumes hybridization proceeds in a two-state all-or-nothing manner. The Gibbs free energy difference between the hybridized and dissociated states can be determined from calculation of NN enthalpy and entropy contributions.
\begin{equation}\label{SIe1}
\Delta G_{NN}^{\circ}(T) = \Delta H_{NN}^{\circ} - T\Delta S_{NN}^{\circ}
\end{equation}
Here,  and  are assumed to be independent of temperature.   and   are determined as the sum of all NN terms in a given configuration. 
\begin{equation}\label{SIe2}
\Delta H_{NN}^{\circ} = \sum_{i}^{n_{NN}}\Delta H_{NN, i}^{\circ} + \sum_{j}^{n_{DE}}\Delta H_{DE, j}^{\circ} + \Delta H_{init}^{\circ}
\end{equation}

\begin{equation}\label{SIe3}
\Delta S_{NN}^{\circ} = \sum_{i}^{n_{NN}}\Delta S_{NN, i}^{\circ} + \sum_{j}^{n_{DE}}\Delta S_{DE, j}^{\circ} + \Delta S_{init}^{\circ} + \Delta S_{sym}^{\circ}
\end{equation}

$\Delta H_{NN, i}^{\circ}$ is the NN enthalpy for a given dinucleotide step within duplex DNA. $\Delta H_{DE, j}^{\circ}$  corresponds to the enthalpic contribution from a given dangling end  base next to the duplex\citep{Santalucia2004TM}. $\Delta H_{init}^{\circ}$ is a duplex initiation term that accounts for terminal effects on duplex enthalpy. $\Delta S_{NN}^{\circ}$  is constructed from analogous NN, dangling end, and initiation terms in addition to a symmetry term, $\Delta S_{sym}^{\circ}$, which applies an entropic penalty to self-complementary sequences for maintaining C2 symmetry. The SantaLucia NN parameters are defined for a solution condition at neutral pH with a sodium concentration of 1M. We use the empirical salt correction developed by Owczarzy\citep{Owczarzy2008PredictingCations} to determine the NN parameters at the sodium concentration used for simulations in this work. The Owczarzy model assumes the affect of counterions on DNA hybridization thermodynamics is purely entropic.

\begin{equation}\label{SIe3}
\Delta S_{NN}^{\circ}(c_{Na^+}) = S_{NN}^{\circ}(1M) + \Delta H_{NN}^{\circ}[(4.29f_{GC}-3.95)\times 10^{-5} \ln{(c_{Na^+})}+ 9.4\times 10^{-6} (\ln{(c_{Na^+})})^2]
\end{equation}

$f_{GC}$ is the fractional G:C content of the duplex and $c_{Na^+}$ is the concentration of sodium counterions in M.

\begin{figure}[ht!]
	\centering
    \includegraphics[width=.85\textwidth]{SIFig6.pdf}
    \caption{Schematic of nearest neighbor contributions for an intact hybridized and out-of-register duplex with dangling ends (DE) stabilization.}
    \label{fig:SIFig4}
\end{figure}

>>>>

% Fig. S1
\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=1.0\textwidth]{FigS1.pdf}
        \caption{Five-fold cross-validation of the SRV VAMP-2 scores to select the optimal number of SRV coordinates for each sequence. A knee in the VAMP-2 plot was identified at the fifth, fourth, third, and second slow modes for AT-all, GC-end, GC-core, and GC-mix, respectively. An embedding of corresponding dimensionality was then used to cluster frames into discrete states. The absence of any significant separation in the training and testing VAMP-2 scores demonstrates that model is not overfitted. \noter{Please re-order panels into standard order with panel labels as per the caption.}}
        \label{fig:SIFig1}
	\end{center}
\end{figure}

@@

% Fig. S2
\begin{figure}[ht!]
	\centering
    \includegraphics[width=0.8\textwidth]{FigS2.pdf}
    \caption{Convergence of the MSM implied time scales $t_i$ as a function of lag time $\tau$. Solid lines indicate maximum likelihood result while dashed lines show the Bayesian ensemble means. The implied time scales for all sequences converge at a lag time of $\tau$ = 1.2 ns (vertical line). The black solid curve marks equality of the implied time scale and lag time and delimits the shaded region wherein the implied time scales are shorter than the lag time and cannot be resolved.}
    \label{fig:SIFig2}
\end{figure}

@@

% Fig. S3
\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=1.0\textwidth]{FigS3.pdf}
        \caption{Chapman-Kolmogorov (CK) tests comparing the probabilities of remaining within each macrostates for each sequence as a function of lag time predicted by an MSM constructed at the $\tau$ = 1.2 ns lag time (dashed blue line) versus those computed from an MSM constructed at the particular lag time (solid black line). The good agreement between these two results provides numerical validation of the Markovian nature of the $\tau$ = 1.2 ns lag time MSM.}
        \label{fig:SIFig3}
	\end{center}
\end{figure}

@@

% Fig. S4
\begin{figure}[ht!]
	\centering
    \includegraphics[width=.75\textwidth]{FigS4.pdf}
    \caption{Exponential fits for both slow (dissociation) and fast (fraying) response for all four sequences during ``computational T-jump'' experiments. From 120 independent 1 $\mu$s simulations, we compiled the slow response data by recording the fraction of sequences with \rood{both OR either} \noter{Pick the right word.} central Watson-Crick base pairs intact as a function of time, and the fast response data as the fraction of sequences with \rood{both OR either} \noter{Pick the right word.} terminal Watson-Crick base pairs intact as a function of time. \noter{Do I have this correct?} We define a Watson-Crick base pair to be intact if the two comlementary bases lie within a linear distance of 1.3 nm. We extracted our computational estimate of $k_d^\mathrm{fast}$ by fitting a decaying exponential to the fraction of bound A:T termini as a function of time $f_\mathrm{unfrayed}(t) = \exp(-k_d^\mathrm{fast}t)$. Similarly, we extracted our computational estimate of $k_d^\mathrm{slow}$ by fitting a decaying exponential to the fraction of hybridized sequences as a function of time $f_\mathrm{hybridized}(t) = \exp(-k_d^\mathrm{slow}t)$.}
    \label{fig:SIFig4}
\end{figure}

@@

% Fig. S5
\begin{figure}[ht!]
	\centering
    \includegraphics[width=0.8\textwidth]{FigS5a.pdf}
\end{figure}
\begin{figure}[ht!]
	\centering
    \includegraphics[width=0.8\textwidth]{FigS5b.pdf}
\end{figure}

@@

The NN model assumes hybridization proceeds in a two-state all-or-nothing manner. The Gibbs free energy difference between the hybridized and dissociated states can be determined from calculation of NN enthalpy and entropy contributions,
\begin{equation}\label{SIe1}
\Delta G_{NN}^{\circ}(T) = \Delta H_{NN}^{\circ} - T\Delta S_{NN}^{\circ}.
\end{equation}
$\Delta H_{NN}^{\circ}$ and $\Delta S_{NN}^{\circ}$ are assumed to be independent of temperature $T$, and are computed from the sum over all NN terms in a given configuration,
\begin{equation}\label{SIe2}
\Delta H_{NN}^{\circ} = \sum_{i}^{n_{NN}}\Delta H_{NN, i}^{\circ} + \sum_{j}^{n_{DE}}\Delta H_{DE, j}^{\circ} + \Delta H_{init}^{\circ},
\end{equation}
\begin{equation}\label{SIe3}
\Delta S_{NN}^{\circ} = \sum_{i}^{n_{NN}}\Delta S_{NN, i}^{\circ} + \sum_{j}^{n_{DE}}\Delta S_{DE, j}^{\circ} + \Delta S_{init}^{\circ} + \Delta S_{sym}^{\circ}.
\end{equation}                          
$\Delta H_{NN, i}^{\circ}$ is the NN enthalpy for a given dinucleotide step within duplex DNA, $\Delta H_{DE, j}^{\circ}$ corresponds to the enthalpic contribution from a given dangling end (DE) base next to the duplex\citep{Santalucia2004TM}, and $\Delta H_{init}^{\circ}$ is a duplex initiation term that accounts for terminal effects on duplex enthalpy. A schematic illustration of the NN and DE contributions for a fully hybridized duplex and an out-of-register shifted state is presented in Fig.~\ref{fig:SIFig4}. $\Delta S_{NN}^{\circ}$ is constructed in an analogous manner from NN, dangling end, and initiation terms, plus a symmetry term, $\Delta S_{sym}^{\circ}$, that applies an entropic penalty to self-complementary sequences for maintaining C2 symmetry.

The SantaLucia NN parameters are defined for a solution condition at neutral pH with a sodium concentration of 1M. We use the empirical salt correction developed by Owczarzy\citep{Owczarzy2008PredictingCations} to determine the NN parameters at the sodium concentration used for simulations in this work. The Owczarzy model assumes the effect of counter ions on DNA hybridization thermodynamics to be purely entropic,
\begin{equation}\label{SIe3}
\Delta S_{NN}^{\circ}(c_{Na^+}) = S_{NN}^{\circ}(1M) + \Delta H_{NN}^{\circ}[(4.29f_{GC}-3.95)\times 10^{-5} \ln{(c_{Na^+})}+ 9.4\times 10^{-6} (\ln{(c_{Na^+})})^2],
\end{equation} 
where $f_{GC}$ is the fractional G:C content of the duplex and $c_{Na^+}$ is the concentration of sodium counter ions measured in M.

\vspace{+0.5in}

\begin{figure}[ht!]
	\centering
    \includegraphics[width=.85\textwidth]{FigS6.pdf}
    \caption{Schematic of nearest neighbor (NN) contributions for a fully hybridized AT-all duplex and an out-of-register duplex with dangling ends (DE).}
    \label{fig:SIFig4}
\end{figure}
